\textcolor{blue}{In-silico molecular transport via perivascular networks in the human intracranial space}

### 2025-06-26 paper/main.tex
Update title
Updates after reading the response.

\title{In-silico molecular enrichment and clearance of the human intracranial space}
\author[1]{Marius Causemann}
\author[1]{Miroslav Kuchta}
\author[2]{Rami Masri}
\author[1,3,*]{Marie E. Rognes }
\affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Division of Applied Mathematics, Brown University, Providence, Rhode Island, USA}
\affil[3]{K. G. Jebsen Centre for Brain Fluid Research, University of Oslo, Norway}
\affil[*]{meg@simula.no}

@@

\subsection*{In-silico predictions of intracranial transport after intrathecal injection}
\begin{figure}
    \centering
    \includegraphics[width=1\textwidth]{figures/figure1.png}
     \caption{
     \textbf{In-silico modelling of transport of the PVS, SAS and parenchyma.}
     A) Illustration of the model geometry including the CSF-filled spaces (ventricles and SAS) $\Omega_{\rm CSF}$, the parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and veins $\Lambda_{\rm vein}$, as well as their interfaces and boundaries; 
     B) 3D rendering of the computational geometry: the parenchyma, CSF (clipped, blue), arterial network (red), and venous network (blue); 
     C) illustration of the intrathecal tracer injection protocol;  
     D) in-silico predictions of tracer concentration after 1, 6, 12, and 24 hours (opacity increasing linearly from 0 to 2 mmol/l); 
     E) sagittal, coronal and axial view of in-silico tracer concentrations overlayed on T1-weighted MR image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red); 
     F) average tracer concentration in each compartment over the first 24\,h; 
     G) total amount of tracer in each compartment over the first 24\,h.}
     \label{fig:results1}
\end{figure}
Using previously published multi-modal magnetic resonance imaging
(MRI) data~\cite{hodneland2019new, deistung2009tof,
  schweser2012quantitative, reichenbach2012future,
  deistung2017overview}, we construct a multiscale computational
representation of the human intracranial compartments consisting of
the CSF spaces and brain parenchyma as three-dimensional (3D) domains
and with the PVSs surrounding major surface arteries and veins as
embedded networks of topologically one-dimensional (1D) curves
(\Cref{fig:results1}A--B). We consider a solute concentration field,
varying in space and time, in the 3D domains and in the PVS networks,
and assume that the solute can cross between these compartments
through semi-permeable membranes. As the drivers and modes of
intracranial transport are under substantial
debate\cite{smith2019going, proulx2021cerebrospinal,
  bohr2022glymphatic, hladky2022glymphatic, betsholtz2024advances},
our first target is to establish a baseline model accounting for a
reasonably conservative set of mechanisms and their integrated effect
over a timescale of several minutes to a few days. To this end, we
assume that the solute will (i) diffuse within all compartments, with
diffusivity depending on the effective properties of the relevant
medium\cite{sykova2008diffusion}; (ii) experience significant
dispersive effects due to the pulsatile flow of CSF induced by the
cardiac and respiratory cycles\cite{vinje2019respiratory,
  sharp2019dispersion, ray2021quantitative, troyetsky2021dispersion};
and (iii) be convected by a (small) net flow of CSF resulting from
production in the choroid plexus with CSF efflux across the upper
convexity\cite{hornkjol2022csf} and from the peristaltic pumping
effect of pulse wave pulsations in surface periarterial
spaces~\cite{mestre2018flow, gjerde2023directional}. Mathematically, this model is represented by a
mixed-dimensional system of coupled time-dependent partial
differential equations\cite{masri2024modelling}, which we solve
numerically with high accuracy using a mass-conserving finite element
scheme and the FEniCS finite element software\cite{alnaes2015fenics,
  kuchta2020assembly} (see Methods). The computational framework and
associated software are all openly available~\cite{ZENODO}.

@@

Now, to examine how CSF pulsatility affects the tracer enrichment patterns,
we consider three variations of the baseline: (i) no CSF production,
(ii) reduced pulsatility and thus decreased dispersion, and (iii) higher pulsatility with increased dispersion. We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion factors by factor of 0.25 and 4, respectively. Note that behaviour directly reflects the nature of our model: The Stokes flow is linear in its inflow boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient.
Without CSF production, transport is considerably delayed
(\Cref{fig:csf}N, O). Even after $12$ hours, tracer remains in the
subtentorial regions around the cerebellum, the brain stem and in the
surrounding CSF. Interestingly, the lack of CSF production instead
allows tracer to travel upwards through the ventricular system,
reaching the third ventricle after around $12$ hours. On the other
hand, if the CSF pulsatility is reduced, we observe rapid transport
towards the upper convexity of the cranium, as in the baseline model,
but the tracer spreads exclusively within the anterior regions
(\Cref{fig:csf}P). This feature can be attributed to the CSF flow
bifurcation posterior to the ambient cistern (\Cref{fig:csf}D):
without sufficient diffusion, the tracer is unable to cross into the
posterior SAS. Indeed, with higher dispersion, the tracer moves
through the quadrigeminal cistern and further upwards into the
longitudinal fissure, with also enrichment of the cerebellum
(\Cref{fig:csf}Q).

>>>>

\title{\textcolor{blue}{In-silico molecular transport via perivascular networks in the human intracranial space}}
\author[1]{Marius Causemann}
\author[1]{Miroslav Kuchta}
\author[2]{Rami Masri}
\author[1,3,*]{Marie E. Rognes }
\affil[1]{Department of Numerical Analysis and Scientific Computing, Simula Research Laboratory, Oslo, Norway}
\affil[2]{Division of Applied Mathematics, Brown University, Providence, Rhode Island, USA}
\affil[3]{K. G. Jebsen Centre for Brain Fluid Research, University of Oslo, Norway}
\affil[*]{meg@simula.no}

@@

\subsection*{In-silico predictions of intracranial transport after intrathecal injection}
\begin{figure}
    \centering
    \includegraphics[width=1\textwidth]{figures/figure1.png}
     \caption{
     \textbf{In-silico modelling of tracer transport in the PVS, SAS and parenchyma.}
     A) Illustration of the model geometry including the CSF-filled spaces (ventricles and SAS) $\Omega_{\rm CSF}$, the parenchyma $\Omega_{\rm PAR}$ and the PVS surrounding arteries $\Lambda_{\rm artery}$ and veins $\Lambda_{\rm vein}$, as well as their interfaces and boundaries; 
     B) 3D rendering of the computational geometry: the parenchyma, CSF (clipped, blue), arterial network (red), and venous network (blue); 
     C) illustration of the intrathecal tracer injection protocol;  
     D) in-silico predictions of tracer concentration after 1, 6, 12, and 24 hours (opacity increasing linearly from 0 to 2 mmol/l); 
     E) sagittal, coronal and axial view of in-silico tracer concentrations overlayed on T1-weighted MR image after 1, 6, 12 and 24\,h (low concentrations transparent, pial surface in pink, arachnoid membrane in cyan, and arteries in dark red); 
     F) average tracer concentration in each compartment over the first 24\,h; 
     G) total amount of tracer in each compartment over the first 24\,h.}
     \label{fig:results1}
\end{figure}
Using previously published multi-modal magnetic resonance imaging
(MRI) data~\cite{hodneland2019new, deistung2009tof,
  schweser2012quantitative, reichenbach2012future,
  deistung2017overview}, we construct a multiscale computational
representation of the human intracranial compartments consisting of
the CSF spaces and brain parenchyma as three-dimensional (3D) domains
and with the PVSs surrounding major surface arteries and veins as
embedded networks of topologically one-dimensional (1D) curves
(\Cref{fig:results1}A--B). We consider a solute concentration field,
varying in space and time, in the 3D domains and in the PVS networks,
and assume that the solute can cross between these compartments
through semi-permeable membranes. As the drivers and modes of
intracranial transport are under substantial
debate\cite{smith2019going, proulx2021cerebrospinal,
  bohr2022glymphatic, hladky2022glymphatic, betsholtz2024advances},
our first target is to establish a baseline model accounting for a
reasonably conservative set of mechanisms and their integrated effect
over a timescale of several minutes to a few days. To this end, we
assume that the solute will (i) diffuse within all compartments, with
diffusivity depending on the effective properties of the relevant
medium\cite{sykova2008diffusion}; (ii) experience significant
dispersive effects due to the pulsatile flow of CSF induced by the
cardiac and respiratory cycles\cite{vinje2019respiratory,
  sharp2019dispersion, ray2021quantitative, troyetsky2021dispersion};
and (iii) be convected by a (small) net flow of CSF resulting from
production in the choroid plexus with CSF efflux across the upper
convexity\cite{hornkjol2022csf} and from the peristaltic pumping
effect of pulse wave pulsations in surface periarterial
spaces~\cite{mestre2018flow, gjerde2023directional}. Mathematically, this model is represented by a
mixed-dimensional system of coupled time-dependent partial
differential equations\cite{masri2024modelling}, which we solve
numerically with high accuracy using a mass-conserving finite element
scheme and the FEniCS finite element software\cite{alnaes2015fenics,
  kuchta2020assembly} (see Methods). The computational framework and
associated software are all openly available~\cite{ZENODO}.

@@

Now, to examine how CSF pulsatility affects the tracer enrichment patterns,
we consider three variations of the baseline: (i) no CSF production,
(ii) reduced pulsatility and thus decreased dispersion, and (iii) higher pulsatility with increased dispersion. We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion factors by 0.25$\times$ and 4$\times$, respectively. Note that behaviour directly reflects the nature of our model: Stokes flow is linear in its inflow boundary conditions, and the diffusion enhancement factor is quadratic in the predicted pressure gradient.
Without CSF production, transport is considerably delayed
(\Cref{fig:csf}N, O). Even after $12$ hours, tracer remains in the
subtentorial regions around the cerebellum, the brain stem and in the
surrounding CSF. Interestingly, the lack of CSF production instead
allows tracer to travel upwards through the ventricular system,
reaching the third ventricle after around $12$ hours. On the other
hand, if the CSF pulsatility is reduced, we observe rapid transport
towards the upper convexity of the cranium, as in the baseline model,
but the tracer spreads exclusively within the anterior regions
(\Cref{fig:csf}P). This feature can be attributed to the CSF flow
bifurcation posterior to the ambient cistern (\Cref{fig:csf}D):
without sufficient diffusion, the tracer is unable to cross into the
posterior SAS. Indeed, with higher dispersion, the tracer moves
through the quadrigeminal cistern and further upwards into the
longitudinal fissure, with also enrichment of the cerebellum
(\Cref{fig:csf}Q).
